On discretely entropy conservative discontinuous Galerkin methods

### 2017-07-30 docs/splitform/main.tex

To the authors knowledge, the construction of entropy conservative and entropy stable schemes have relied on the use of lumped diagonal mass matrices and nodal collocation (diagonal norm SBP operators).  Entropy stable high order finite element and DG methods have been proposed \cite{hughes1986new}, but the proofs are often given at the continuous level, relying on exact integration or the chain rule, neither of which holds at the discrete level.  Thus, while DG-SEM based on flux differencing is entropy conservative, this is tied to a specific choice of basis and quadrature.  Diagonal-norm SBP operators can be constructed for triangles and tetrahedra \cite{chin1999higher,  hicken2016multidimensional, chen2017entropy}; however, the number of nodal points for such operators is greater than the dimension of the natural polynomial approximation space.  Furthermore, appropriate point sets have only been constructed for $N \leq 4$ in three dimensions \cite{zhebel2014comparison}.  
%Furthermore, diagonal-norm SBP operators do not naturally generalize to triangles or tetrahedra \cite{helenbrook2009existence}.

This work focuses on the construction of entropy conservative high order DG schemes for systems of conservation laws.  In order to generalize beyond collocation DG and diagonal norm SBP operators, we will consider dense (non-lumped) mass matrices and more general quadrature rules (e.g.\ rules without boundary points, rules with more points than the dimension of the corresponding approximation space).  We will present proofs of discrete entropy stability in terms of integrals and $L^2$ projections, which we assume are computed using an appropriate quadrature rule.  We stress that, while we utilize a continuous notation involving integrals rather than discrete sums, the proofs rely only on properties of quadrature-based integration and $L^2$ projection.

The outline of the paper is as follows: Section~\ref{sec:intro} will briefly review the construction of entropy conservative collocation methods on a single element.  Section~\ref{sec:ecdg} will describe how to construct analogous entropy conservative methods on single element.  Section~\ref{sec:ecdg2} will discuss extensions to multiple elements, including comparisons of different coupling terms and entropy stable fluxes.  We note that, for simplicity and clarity of presentation, the proposed methods and proofs are given in one dimension.  However, the extension to multiple dimensions (including simplicial elements) is straightforward and will be described in an upcoming manuscript.  
Finally, Section~\ref{sec:num} presents numerical results which verify the high order accuracy and discrete entropy conservation of the proposed methods.

@@

\section{Entropy conservative diagonal norm collocation DG methods}
\label{sec:intro}

We will begin by reviewing continuous entropy theory, as well as existing high order methods which are provably entropy conservative at the discrete level.  These methods were introduced in \cite{fisher2013high, carpenter2014entropy} for tensor product elements, though we adopt methods of proof introduced more recently in \cite{gassner2017br1, chen2017entropy}.  These methods all share a basis in diagonal norm summation by parts operators.  Because they can be derived from a nodal discontinuous Galerkin framework through the collocation of nodal points and quadrature points, we refer to them as ``diagonal norm collocation DG methods'' in this work.

@@

We assume now that the domain is the interval $[-1,1]$.  Integrating the entropy inequality over this interval and using the definition of the entropy potential then yields
\begin{equation}
\int_{-1}^1 \pd{U(\bm{u})}{t} + \left.\LRp{\bm{v}^T\bm{f}(\bm{u}(\bm{v}))-\psi(\bm{v})}\right|_{-1}^1 \leq 0.  
\label{eq:consentropy}
\end{equation}
The focus of this work is the construction of methods which satisfy a discrete analogue of this entropy inequality.  We begin by reviewing existing discretely entropy conservative methods based on collocation and diagonal norm summation by parts (SBP) operators.

\subsection{Entropy conservative collocation methods based on SBP operators}

@@

This formulation conserves the quantities $\bm{u}_i$, but (except in specific cases) does not satisfy a discrete analogue of the statement of entropy conservation (\ref{eq:consentropy}).  Entropy conservative schemes may be constructed based on the SBP property and a ``flux differencing'' approach \cite{gassner2016split} involving a definition of entropy conservative/stable fluxes given by Tadmor \cite{tadmor1987numerical, tadmor2003entropy}.  
\begin{definition}
Let $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ be a bivariate function such that it is symmetric and consistent with the flux function $\bm{f}(\bm{u})$
\[
\bm{f}_S(\bm{u}_L,\bm{u}_R) = \bm{f}_S(\bm{u}_R,\bm{u}_L), \qquad \bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u})
\]
The numerical flux $\bm{f}_S(\bm{u}_L, \bm{u}_R)$ is entropy conservative if, for entropy variables $\bm{v}_L = \bm{v}(\bm{u}_L), \bm{v}_R = \bm{v}(\bm{u}_R)$
\[
\LRp{\bm{v}_L - \bm{v}_R}^T \bm{f}_S(\bm{u}_L,\bm{u}_R) = (\psi_L - \psi_R), \qquad \psi_L = \psi(\bm{v}(\bm{u}_L)), \quad \psi_R = \psi(\bm{v}(\bm{u}_R)).  
\]
Similarly, a flux $\bm{f}_S(\bm{u}_L, \bm{u}_R)$ is referred to as entropy stable if $\LRp{\bm{v}_L - \bm{v}_R}^T \bm{f}_S(\bm{u}_L,\bm{u}_R) \leq (\psi_L - \psi_R)$.
\label{def:tadmor}
\end{definition}
An entropy conservative nodal collocation method can then be derived by modifying the flux derivative as follows
\begin{equation}
\pd{\bm{u}_i}{t} + \sum_{j=1}^{N+1}\LRp{\LRp{2\bm{D}\otimes \bm{I}}_{ij}\bm{f}_S(\bm{u}_i,\bm{u}_j) +\LRp{\bm{M}^{-1}\bm{B}}_{ij}(\bm{f}_i^* - \bm{f}(\bm{u}_i))}= 0.
\label{eq:dgsem}
\end{equation}
The replacement of the derivative of the flux function with flux differencing allows one to prove that (\ref{eq:dgsem}) conserves a discrete entropy \cite{gassner2017br1,chen2017entropy}.  We reproduce these proofs here, as we will refer to them in following sections.  
\begin{theorem}
The collocation method (\ref{eq:dgsem}) is discretely entropy conservative in the sense that
\[
\bm{1}^T\LRp{\bm{M}\pd{U(\bm{u})}{t} + \bm{B}\LRp{\bm{v}^T\bm{f}^* - \bm{\psi}}} = 0,
\]
where $\bm{1}$ denotes the vector of ones and $\bm{\psi}_i = \psi(\bm{v}_i)$ is the entropy potential in terms of the nodal value of $\bm{u}_i$.  
\label{thm:dgsem}
\end{theorem}
\begin{proof}
Let $\bm{v}_i = \bm{v}(\bm{u}_i)$ denote the entropy variables, defined in terms of the nodal values $\bm{u}_i$.  We multiply (\ref{eq:dgsem}) by $\LRp{\bm{M}\bm{v}}^T$.  
Using the fact that $\bm{M}$ is diagonal, the time derivative term can be manipulated to 
\[
\bm{v}^T\bm{M}\pd{\bm{u}}{t} = \sum_{i=1}^{N+1} \bm{M}_{ii} \bm{v}_i^T \pd{\bm{u}_i}{t} = \sum_{i=1}^{N+1} \bm{M}_{ii} \pd{S_i}{t} = \bm{1}^T\bm{M}\pd{S_i}{t},
\]
where we have used continuity in time to arrive at the point-wise relation 
\[
\bm{v}^T \pd{\bm{u}}{t} = \pd{U(\bm{u})}{\bm{u}}^T \pd{\bm{u}}{t} = \pd{U(\bm{u})}{t}.
\]

@@

\note{Fix notation: switch $V_h$ to $P^N$}
We assume that the domain $\Omega \in \mathbb{R}^d$ is exactly triangulated by a mesh which consists of the union of non-overlapping elements $D^k$.  We further assume that each element $D^k$ is the image of a reference element $\widehat{D}$ under the local elemental mapping 
\[
x^k = \bm{\Phi}^k \widehat{x},
\]
where $x^k$ denote physical coordinates on $D^k$ and $\widehat{x}$ denote coordinates on the reference element.  Finally, we identify $J$ as the determinant of the Jacobian of $\bm{\Phi}^k$, and refer to it as the Jacobian for the remainder of this work.  For clarity and simplicity of presentation, we assume $\bm{\Phi}^k$ is affine in this case, such that $J$ is constant over each element $D^k$.

We will approximate solution components over each element $D^k$ from an approximation space $V_h\LRp{D^k}$, which we define as the composition of the mapping $\bm{\Phi}^k$ and a reference approximation space $V_h\LRp{\widehat{D}}$
\[
V_h\LRp{D^k} = \bm{\Phi}^k \circ V_h\LRp{\widehat{D}}.
\]
The global approximation space $V_h\LRp{\Omega}$ is then defined as the direct sum of local approximation spaces
\[
V_h\LRp{\Omega} = \bigoplus_{D^k}V_h\LRp{D^k}.  
\]
We assume that $V_h(D^k) \subset H^1(D^k)$; in other words, that all elements of each local approximation space are square integrable and admit well-defined traces.

We define the $L^2$ norm and inner products over the element $D^k$ and the surface of the element $\partial D^k$\footnote{In one dimension, $D^k$ is some interval $[x_L,x_R]$ and the $L^2$ inner product over the surface reduces to 
\[
\LRa{u,v}_{\partial D^k} = u(x_L)v(x_L) + u(x_R)v(x_R).
\]
However, the presented proofs in this work will not utilize this simplification, and are thus generalizable to higher dimensions in a straightforward manner.}
\[
\LRp{u,v}_{D^k} = \int_{D^k} u v\diff{x}, \qquad \nor{u}^2_{D^k} = (u,u)_{D^k}, \qquad \LRa{u,v}_{\partial D^k} = \int_{\partial D^k} u v \diff{x}.
\]
We also define the local $L^2$ projection operator $\Pi_N: L^2(D^k)\rightarrow V_h(D^k)$ such that
\[
\LRp{\Pi_N u,v}_{D^k} = \LRp{u,v}_{D^k}, \qquad \forall v\in V_h\LRp{D^k}.  
\]
Throughout this work, we will use that this definition of the $L^2$ projection holds even when integrals are approximated using quadrature.  We will make no assumptions on quadrature except that \note{specify assumptions - integrate polynomials of degree $N$ on the reference element exactly, such that integration by parts holds.  Add footnote that we can also generalize to the under-integrated case.  }

Finally, we define the jump and average of discontinuous functions across element interfaces.  Let $D^{k,+}$ denote the neighboring element of across a face $f$ of $D^k$, and let $u^+$ denote the values of $u$ on $D^{k,+}$.  The jump of $u$ across $f$ on $D^k$ is then defined as
\[
\jump{u} = u^+ - u, \qquad \avg{u} = \frac{u^+ + u}{2}.
\]
The jump and average of vector fields are defined component-wise using the jumps and averages of components.  When the face $f$ lies on the boundary, we define $u^+ = u$, such that $\jump{u} = 0$ and $\avg{u} = u$ on $\partial \Omega$.

@@

It is well known that DG formulations based on a split formulation can be shown to conserve the square entropy $u^2$ \cite{gassner2013skew, ranocha2017extended}.  This form may be also be recovered using flux differencing under the two-point flux
\[
f_S(u_L,u_R) = \frac{1}{6}(u_L^2 + u_Lu_R + u_R^2).
\]
Applying (\ref{eq:fluxdiff}) yields the split form of Burgers' equation
\[
2\left.\pd{f_S(u(x),u(y))}{x}\right|_{y=x} = \frac{1}{3}\left.\pd{\LRp{u(x)^2 + u(x)u(y) + u(y)^2}}{x}\right|_{y=x} = \frac{1}{3}\LRp{\pd{u^2}{x} + u\pd{u}{x}}.
\]
We note that recovering the split-form of Burgers' equation relies on the property
%\footnote{This is also the reason why the factor of $2$ is necessary when using flux differentiation (\ref{eq:fluxdiff}); the differentiation of terms which depend only on $y$ vanish, \note{finish...}.}
that 
\[
\pd{\LRp{u(y)^2}}{x} = u(y)^2\pd{\LRp{1}}{x} = 0.  
\]
Thus, recovery of the split form can be achieved by using flux differencing in combination with any differential operator $D$ such that $D1 = 0$.  
%\[
%D:L^2 \rightarrow V_h, \qquad D1 = 0.
%\]
%For Burgers' equation (as well as any skew-symmetric splitting), replacing the exact derivative $\pd{}{x}$ by $D$ will again recover a statement of the split formulation (\ref{eq:burgerssplit}) 
%\[
%\pd{u}{t} + \frac{1}{3}\LRp{Du^2 + uDu} = 0.
%\]
%The introduction of $D$ is necessary when adapting the flux differentiation procedure in the context of numerical discretizations.

@@

Both the conservative and split forms of Burgers' equation are equivalent at the continuous level (assuming continuity of $u$ and exact differentiation); however, this is no longer true at the discrete level.  Suppose now that the solution is approximated by $u \in V_h(\Omega)$ on a single element.  Since $f(u) = u^2/2 \not\in V_h(\Omega)$ for general choices of $u, V_h$, it is not possible to differentiate $f(u)$ exactly without prior knowledge of the form of the nonlinearity.  Indeed, it is more common to first project $f(u)$ to $V_h$, then differentiate, such that the discrete scheme is equivalent to solving for $u \in V_h(\Omega)$ such that
\[
\int_{-1}^1 \LRp{\pd{u}{t} + \pd{\Pi_N f(u)}{x}}w \diff{x}  
%\int_{-1}^1 \LRp{\pd{u}{t} +D_N f(u)}w \diff{x} 
=0, \qquad \forall w \in V_h(\Omega).  
\]
This scheme does not, in general, conserve energy.  %Moreover, if $\Pi_N$ is computed using quadrature, rewriting the scheme into this form suggests that de-aliasing techniques based on over-integration hit a ``bottleneck'', due to the fact that increasing quadrature accuracy only increases the accuracy of the projection \cite{kirby2003aliasing}.

We can derive a semi-discrete Galerkin scheme based which conserves energy discretely using flux differencing and a different choice of discrete differentiation operator $D_N$, which was motivated by the SBP operators introduced in \cite{chen2017entropy}.  Let $w(x)$ be some bounded real-valued function; then, we can define the discrete differential operator $D_N: H^1\rightarrow V_h$ such that, for $u\in H^1$, 
\begin{align}
\LRp{D_N u, vw}_{\widehat{D}} = \LRp{\pd{ \Pi_N u}{x},vw}_{\widehat{D}} + \frac{1}{2}{\LRa{{u - \Pi_N u}, vw\bm{n}_x}}_{\partial \widehat{D}} + \frac{1}{2}\LRa{u - \Pi_Nu ,\Pi_N\LRp{vw}\bm{n}_x }_{\widehat{D}}, \qquad \forall v\in V_h.
\label{eq:dgd}
\end{align}
This operator satisfies a form of integration by parts, which is summarized by the following lemma:
\begin{lemma}
\label{lemma:dgd_local}
The linear operator $D_N$ satisfies a ``weighted'' integration by parts property with respect the $L^2$ inner product and any bounded real-valued function $w(x)$ 
\[
\LRp{D_N u, wv}_{\widehat{D}} = \LRa{u,wv\bm{n}_x}_{\partial \widehat{D}} - \LRp{u,D_N (w v)}_{\widehat{D}}.
\]
\end{lemma}
\begin{proof}
Integrating the left hand side by parts gives
\[
-\LRp{\pd{ \Pi_N (vw)}{x},u}_{\widehat{D}} + \frac{1}{2}{\LRa{{u - \Pi_N u}, vw\bm{n}_x}}_{\partial \widehat{D}} + \frac{1}{2}\LRa{u + \Pi_Nu ,\Pi_N\LRp{vw}\bm{n}_x }_{\partial \widehat{D}}, \qquad \forall v\in V_h.
\]
Rearranging boundary terms yields
\[
-\LRp{\pd{ \Pi_N (vw)}{x},u}_{\widehat{D}} + \frac{1}{2}\LRa{vw + \Pi_N(vw) , u\bm{n}_x }_{\partial \widehat{D}}  - \frac{1}{2}\LRa{vw - \Pi_N (vw), \Pi_N u \bm{n}_x}_{\partial \widehat{D}}.  
\]
Adding and subtracting $\LRa{vw,u\bm{n}_x}_{\partial \widehat{D}}$ completes the proof.
\end{proof}

>>>>

To the authors knowledge, the construction of entropy conservative and entropy stable schemes have relied on the use of lumped diagonal mass matrices and nodal collocation.  Since these can be reinterpreted as diagonal norm SBP operators, we refer to DG methods with these properties as diagonal norm SBP-DG methods.  Entropy stable high order finite element and DG methods which do not fall under the diagonal norm SBP-DG category have been proposed \cite{hughes1986new}, but the proofs are often given at the continuous level, relying on exact integration or the chain rule, neither of which holds at the discrete level.

On the other hand, while entropy conservative schemes can be constructed by combining flux differencing with diagonal norm SBP operators, this can restrict one to specific choices of basis and quadrature.  Appropriate diagonal norm SBP operators can be constructed on tensor product elements based on a DG-SEM discretization.  Diagonal-norm SBP operators can also be constructed for triangles and tetrahedra \cite{chin1999higher,  hicken2016multidimensional, chen2017entropy}; however, the number of nodal points for such operators is greater than the dimension of the natural polynomial approximation space.  Furthermore, appropriate point sets have only been constructed for $N \leq 4$ in three dimensions \cite{zhebel2014comparison}, and (to the author's knowledge) the construction of diagonal norm SBP-DG methods has not yet been performed for less standard elements such as pyramids \cite{chan2016orthogonal}.  
%Furthermore, diagonal-norm SBP operators do not naturally generalize to triangles or tetrahedra \cite{helenbrook2009existence}.

This work focuses on the construction of entropy conservative high order DG schemes for systems of conservation laws.  In order to generalize beyond diagonal norm SBP-DG methods, we will consider dense (non-lumped) mass matrices and more general quadrature rules (e.g.\ rules without boundary points, rules with more points than the dimension of the corresponding approximation space).  We will present proofs of discrete entropy stability in terms of integrals and $L^2$ projections, which we assume are computed using an appropriate quadrature rule.  We stress that, while we utilize a continuous notation involving integrals rather than discrete sums, the proofs rely only on properties of quadrature-based integration and $L^2$ projection.

The outline of the paper is as follows: Section~\ref{sec:intro} will briefly review the construction of entropy conservative diagonal norm SBP-DG methods on a single element.  Section~\ref{sec:ecdg} will describe how to construct analogous entropy conservative methods on single element.  Section~\ref{sec:ecdg2} will discuss extensions to multiple elements, including comparisons of different coupling terms and entropy stable fluxes.  We note that, for simplicity and clarity of presentation, the proposed methods and proofs are given in one dimension.  However, the extension to multiple dimensions (including simplicial elements) is straightforward and will be described in an upcoming manuscript.  
Finally, Section~\ref{sec:num} presents numerical results which verify the high order accuracy and discrete entropy conservation of the proposed methods.

@@

\section{Entropy conservative diagonal norm SBP-DG methods}
\label{sec:intro}

We will begin by reviewing continuous entropy theory and existing high order methods which are provably entropy conservative at the discrete level.  These methods were introduced in \cite{fisher2013high, carpenter2014entropy} for tensor product elements, though we adopt methods of proof introduced more recently in \cite{gassner2017br1, chen2017entropy}.  These methods all share a basis in diagonal norm summation by parts operators, and can be derived from a nodal discontinuous Galerkin framework through the collocation of nodal points and quadrature points.

@@

We assume now that the domain is the interval $[-1,1]$.  Integrating the entropy inequality over this interval and using the definition of the entropy potential then yields
\begin{equation}
\int_{-1}^1 \pd{U(\bm{u})}{t} + \left.\LRp{\bm{v}^T\bm{f}(\bm{u}(\bm{v}))-\psi(\bm{v})}\right|_{-1}^1 \leq 0.  
\label{eq:consentropy}
\end{equation}
The focus of this work is the construction of methods which satisfy a discrete analogue of this entropy inequality.  We begin by reviewing existing discretely entropy conservative methods based on diagonal norm summation by parts (SBP) operators.

\subsection{Entropy conservative methods based on diagonal-norm SBP operators}

@@

This formulation conserves the quantities $\bm{u}_i$, but (except in specific cases) does not satisfy a discrete analogue of the statement of entropy conservation (\ref{eq:consentropy}).  Entropy conservative schemes may be constructed based on the SBP property and a ``flux differencing'' approach \cite{gassner2016split} involving a definition of entropy conservative/stable fluxes given by Tadmor \cite{tadmor1987numerical, tadmor2003entropy}.  
\begin{definition}
Let $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ be a bivariate function such that it is symmetric and consistent with the flux function $\bm{f}(\bm{u})$
\[
\bm{f}_S(\bm{u}_L,\bm{u}_R) = \bm{f}_S(\bm{u}_R,\bm{u}_L), \qquad \bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u})
\]
The numerical flux $\bm{f}_S(\bm{u}_L, \bm{u}_R)$ is entropy conservative if, for entropy variables $\bm{v}_L = \bm{v}(\bm{u}_L), \bm{v}_R = \bm{v}(\bm{u}_R)$
\[
\LRp{\bm{v}_L - \bm{v}_R}^T \bm{f}_S(\bm{u}_L,\bm{u}_R) = (\psi_L - \psi_R), \qquad \psi_L = \psi(\bm{v}(\bm{u}_L)), \quad \psi_R = \psi(\bm{v}(\bm{u}_R)).  
\]
Similarly, a flux $\bm{f}_S(\bm{u}_L, \bm{u}_R)$ is referred to as entropy stable if $\LRp{\bm{v}_L - \bm{v}_R}^T \bm{f}_S(\bm{u}_L,\bm{u}_R) \leq (\psi_L - \psi_R)$.
\label{def:tadmor}
\end{definition}
An entropy conservative diagonal norm SBP-DG method can then be derived by modifying the flux derivative as follows
\begin{equation}
\pd{\bm{u}_i}{t} + \sum_{j=1}^{N+1}\LRp{\LRp{2\bm{D}\otimes \bm{I}}_{ij}\bm{f}_S(\bm{u}_i,\bm{u}_j) +\LRp{\bm{M}^{-1}\bm{B}}_{ij}(\bm{f}_i^* - \bm{f}(\bm{u}_i))}= 0.
\label{eq:dgsem}
\end{equation}
The replacement of the derivative of the flux function with flux differencing allows one to prove that (\ref{eq:dgsem}) conserves a discrete entropy \cite{gassner2017br1,chen2017entropy}.  We reproduce these proofs here, as we will refer to them in following sections.  
\begin{theorem}
The nodal collocation method (\ref{eq:dgsem}) is discretely entropy conservative in the sense that
\[
\bm{1}^T\LRp{\bm{M}\pd{U(\bm{u})}{t} + \bm{B}\LRp{\bm{v}^T\bm{f}^* - \bm{\psi}}} = 0,
\]
where $\bm{1}$ denotes the vector of ones and $\bm{\psi}_i = \psi(\bm{v}_i)$ is the entropy potential in terms of the nodal value of $\bm{u}_i$.  
\label{thm:dgsem}
\end{theorem}
\begin{proof}
Let $\bm{v}_i = \bm{v}(\bm{u}_i)$ denote the entropy variables, defined in terms of the nodal values $\bm{u}_i$.  We multiply (\ref{eq:dgsem}) by $\LRp{\bm{M}\bm{v}}^T$.  
Using the fact that $\bm{M}$ is diagonal, the time derivative term can be manipulated to 
\[
\bm{v}^T\bm{M}\pd{\bm{u}}{t} = \sum_{i=1}^{N+1} \bm{M}_{ii} \bm{v}_i^T \pd{\bm{u}_i}{t} = \sum_{i=1}^{N+1} \bm{M}_{ii} \pd{S_i}{t} = \bm{1}^T\bm{M}\pd{S_i}{t},
\]
where we have used continuity in time to arrive at the point-wise relation 
\[
\bm{v}^T \pd{\bm{u}}{t} = \pd{U(\bm{u})}{\bm{u}}^T \pd{\bm{u}}{t} = \pd{U(\bm{u})}{t}.
\]

@@

%\note{Fix}

We assume that the domain $\Omega$ is an interval, which is decomposed into non-overlapping intervals (elements) $D^k$ of size $h$.  We define the reference element $\widehat{D} = [-1,1]$, such that $D^k$ is the image of $\widehat{D}$ under an affine mapping $x = \Phi^k(\widehat{x})$.  We approximate the solution over the reference element using degree $N$ polynomials 
\[
P^N\LRp{\widehat{D}} = \LRc{\widehat{x}^i, \quad \widehat{x} \in \widehat{D}, \quad 0\leq i \leq N}.
\]
%We define an approximation space over the reference element $\widehat{D}$ as the space of degree $N$ polynomials 
%\[
%P^N\LRp{\widehat{D}} = \LRc{\widehat{x}^i, \qquad 0\leq i \leq N}.
%\]
%We will approximate solution components over each element $D^k$ using 
% $V_h\LRp{D^k}$, which we define as the composition of the mapping $\bm{\Phi}^k$ and $P^N\LRp{\widehat{D}}$
%\[
%V_h\LRp{D^k} = \bm{\Phi}^k \circ V_h\LRp{\widehat{D}}.
%\]
We approximate the solution over a physical element by mapping $P^N\LRp{\widehat{D}}$ to $D^k$ under $\Phi^k$, and define the global approximation space $V_h$ as the direct sum of local approximation spaces
\[
P^N\LRp{D^k} =  \bm{\Phi}^k \circ P^N\LRp{\widehat{D}}, \qquad V_h = \bigoplus_{k}P^N\LRp{D^k}.
\]
We define the $L^2$ norm and inner products over the element $D^k$ and the surface of the element $\partial D^k$
\[
\LRp{u,v}_{D^k} = \int_{D^k} u v\diff{x} =  \int_{D^k} u v J^k \diff{\widehat{x}}, \qquad \nor{u}^2_{D^k} = (u,u)_{D^k}, \qquad \LRa{u,v}_{\partial D^k} = \int_{\partial D^k} u v \diff{x},
\]
where $J^k$ is the determinant of the Jacobian of $\Phi^k$.  
We also define the local $L^2$ projection operator $\Pi_N: L^2(D^k)\rightarrow P^N(D^k)$ such that
\[
\LRp{\Pi_N u,v}_{D^k} = \LRp{u,v}_{D^k}, \qquad \forall v\in V_h\LRp{D^k}.  
\]
In one dimension, $D^k$ is some interval and the $L^2$ inner product over the surface reduces to 
\[
\LRa{u,v}_{\partial D^k} = u(x_L)v(x_L) + u(x_R)v(x_R).
\]  
However, the presented proofs in this work will not directly utilize this simplification in order to remain generalizable to higher dimensions in a straightforward manner.

Throughout this work, we will assume that integrals and inner products are approximated using some quadrature rule consisting of $N_q$ points $\widehat{x}_i$ and weights $w_i$ over the reference element $\widehat{D}$.  
\[
\LRp{u,v}_{D^k} \coloneqq \sum_{i=1}^{N_q} u(\widehat{x}_i)v(\widehat{x}_i)w_i J^k.     
\]
This also defines a quadrature-based $L^2$ projection operator.  We emphasize that, throughout this work, we will only use properties of inner products and projection operators which hold when integrals are approximated using quadrature.  We assume only that the quadrature is sufficiently accurate to integrate degree $(2N-1)$ polynomials, such that integration by parts holds on the reference element.  However, we will also allow for quadrature with an arbitrary number of points and arbitrary accuracy.

%\note{specify assumptions - integrate polynomials of degree $N$ on the reference element exactly, such that integration by parts holds.  Add footnote that we can also generalize to the under-integrated case.  }

Finally, we define the jump and average of discontinuous functions across element interfaces.  Let $D^{k,+}$ denote the neighboring element of across a face $f$ of $D^k$, and let $u^+$ denote the values of $u$ on $D^{k,+}$.  The jump of $u$ across $f$ on $D^k$ is then defined as
\[
\jump{u} = u^+ - u, \qquad \avg{u} = \frac{u^+ + u}{2}.
\]
The jump and average of vector fields are defined component-wise using the jumps and averages of components.  When the face $f$ lies on the boundary, we define $u^+ = u$ such that $\jump{u} = 0$ and $\avg{u} = u$ on $\partial \Omega$.

@@

It is well known that DG formulations based on a split formulation can be shown to conserve the square entropy $u^2$ \cite{gassner2013skew, ranocha2017extended}.  This form may be also be recovered using flux differencing under the two-point flux
\[
f_S(u_L,u_R) = \frac{1}{6}(u_L^2 + u_Lu_R + u_R^2).
\]
Applying (\ref{eq:fluxdiff}) yields the split form of Burgers' equation
\[
2\left.\pd{f_S(u(x),u(y))}{x}\right|_{y=x} = \frac{1}{3}\left.\pd{\LRp{u(x)^2 + u(x)u(y) + u(y)^2}}{x}\right|_{y=x} = \frac{1}{3}\LRp{\pd{u^2}{x} + u\pd{u}{x}}.
\]
We note that recovering the split-form of Burgers' equation relies on the property
%\footnote{This is also the reason why the factor of $2$ is necessary when using flux differentiation (\ref{eq:fluxdiff}); the differentiation of terms which depend only on $y$ vanish, \note{finish...}.}
that 
\[
\pd{\LRp{u(y)^2}}{x} = u(y)^2\pd{\LRp{1}}{x} = 0.  
\]
Thus, recovery of the split form can be achieved by using flux differencing in combination with any differential operator $D$ such that $D1 = 0$.

@@

Both the conservative and split forms of Burgers' equation are equivalent at the continuous level (assuming continuity of $u$ and exact differentiation); however, this is no longer true at the discrete level.  Suppose now that the solution is approximated by $u \in P^N\LRp{\widehat{D}}$ on the reference element.  Since $f(u) = u^2/2 \not\in P^N\LRp{\widehat{D}}$ for general choices of $u$, it is not possible to differentiate $f(u)$ exactly without prior knowledge of the form of the nonlinearity.  Indeed, it is more common to first project $f(u)$ to $P^N\LRp{\widehat{D}}$, then differentiate, such that the discrete scheme is equivalent to solving for $u \in P^N\LRp{\widehat{D}}$ such that
\[
\int_{-1}^1 \LRp{\pd{u}{t} + \pd{\Pi_N f(u)}{x}}w \diff{x}  
%\int_{-1}^1 \LRp{\pd{u}{t} +D_N f(u)}w \diff{x} 
=0, \qquad \forall w \in P^N\LRp{\widehat{D}}.  
\]
This scheme does not, in general, conserve energy.  %Moreover, if $\Pi_N$ is computed using quadrature, rewriting the scheme into this form suggests that de-aliasing techniques based on over-integration hit a ``bottleneck'', due to the fact that increasing quadrature accuracy only increases the accuracy of the projection \cite{kirby2003aliasing}.

We can derive a semi-discrete Galerkin scheme based which conserves energy discretely using flux differencing and a different choice of discrete differentiation operator $D_N$, which was motivated by the SBP operators introduced in \cite{chen2017entropy}.  Define $H^1\LRp{\widehat{D}}$ as the Sobolev space
\[
\nor{u}_{H^1\LRp{\widehat{D}}}^2 = \nor{u}_{L^2\LRp{\widehat{D}}}^2 + \nor{\pd{u}{x}}_{L^2\LRp{\widehat{D}}}^2, \qquad H^1\LRp{\widehat{D}} = \LRc{ u \in L^2\LRp{\widehat{D}}, \qquad \nor{u}_{H^1\LRp{\widehat{D}}} < \infty}.
\]
Because functions $u \in H^1\LRp{\widehat{D}}$ admit a well-defined trace \cite{brenner2007mathematical}, we can define the following discrete differential operator as a map from $H^1$ to $P^N$:
\begin{definition}
Let $w(x)$ be a bounded function on $\widehat{D}$.  Define $D_N: H^1\LRp{\widehat{D}}\rightarrow P^N\LRp{\widehat{D}}$ such that, for $u\in H^1\LRp{\widehat{D}}$, 
\begin{align}
&\LRp{D_N u, vw}_{\widehat{D}} = \LRp{\pd{ \Pi_N u}{x},vw}_{\widehat{D}}\nonumber \\
&+ \frac{1}{2}{\LRa{{u - \Pi_N u}, vw\bm{n}_x}}_{\partial \widehat{D}} + \frac{1}{2}\LRa{u - \Pi_Nu ,\Pi_N\LRp{vw}\bm{n}_x }_{\widehat{D}}, \qquad \forall v\in P^N\LRp{\widehat{D}}.
\label{eq:dgd}
\end{align}
\end{definition}
This operator satisfies a form of integration by parts, which is summarized by the following lemma:
\begin{lemma}
\label{lemma:dgd_local}
The linear operator $D_N$ satisfies a ``weighted'' integration by parts property with respect the $L^2$ inner product and any bounded real-valued function $w(x)$ 
\[
\LRp{D_N u, wv}_{\widehat{D}} = \LRa{u,wv\bm{n}_x}_{\partial \widehat{D}} - \LRp{u,D_N (w v)}_{\widehat{D}}.
\]
\end{lemma}
\begin{proof}
Integrating the left hand side by parts gives
\[
-\LRp{\pd{ \Pi_N (vw)}{x},u}_{\widehat{D}} + \frac{1}{2}{\LRa{{u - \Pi_N u}, vw\bm{n}_x}}_{\partial \widehat{D}} + \frac{1}{2}\LRa{u + \Pi_Nu ,\Pi_N\LRp{vw}\bm{n}_x }_{\partial \widehat{D}}, \qquad \forall v\in P^N\LRp{\widehat{D}}.
\]
Rearranging boundary terms yields
\[
-\LRp{\pd{ \Pi_N (vw)}{x},u}_{\widehat{D}} + \frac{1}{2}\LRa{vw + \Pi_N(vw) , u\bm{n}_x }_{\partial \widehat{D}}  - \frac{1}{2}\LRa{vw - \Pi_N (vw), \Pi_N u \bm{n}_x}_{\partial \widehat{D}}.  
\]
Adding and subtracting $\LRa{vw,u\bm{n}_x}_{\partial \widehat{D}}$ completes the proof.
\end{proof}

Direct computations also show that $\LRp{D_N 1,vw}_{\widehat{D}} = 0$ for all bounded weights $w(x)$ and $\forall v \in P^N\LRp{\widehat{D}}$.  Thus, we can apply (\ref{eq:fluxdiff}) with $D_N$ to construct a discrete formulation for Burgers' equation
\begin{align}
\LRp{\pd{u}{t},v} + \frac{1}{3}\LRp{D_N u^2 + uD_Nu,v} &= \LRp{\pd{u}{t},v}_{\widehat{D}} + \frac{1}{3}\LRp{D_N u^2,v}_{\widehat{D}} + \frac{1}{3}\LRp{D_Nu,uv}_{\widehat{D}},
%\LRp{\pd{u}{t},v}_{\widehat{D}} + \frac{1}{3}\LRp{\pd{\Pi_N u^2}{x},v}_{\widehat{D}} + \frac{1}{3}\LRp{D_Nu,uv}_{\widehat{D}}
\qquad \forall v\in P^N\LRp{\widehat{D}}.
\label{eq:burgersdiscrete}
\end{align}
To show this formulation is discretely energy conservative, we take $v=u$ and use Lemma~\ref{lemma:dgd_local} to cancel volume terms, yielding
\[
\frac{1}{2}\pd{}{t}\LRp{u,u}_{\widehat{D}} = -\frac{1}{3}\LRa{u^2,u}_{\partial \widehat{D}},
\]
which is a discrete analogue of the conservation condition (\ref{eq:burgersconservation}).  This is equivalent to the skew-symmetric split formulation described in \cite{gassner2013skew}; however, as noted in \cite{chen2017entropy}, it can be derived from flux differencing without explicit knowledge of the skew form.  We will show later in Section~\ref{sec:recovery} that a natural extension of this formulation to multiple elements recovers the weak form of the scheme introduced in \cite{ranocha2017extended}.
